\mytitle{}

### 2018-08-08 paper/paper.tex
More ideas + refs

?? What makes RL a suitable method for network anomaly detection, what features are most relevant?

@@

To test this, we configured the network topology using $k=2$ teams, $l=3$ intermediate nodes per team, $m=2$ agents per intermediate node, and $n \in \{2, 4, 8, 16\}$ hosts per learner.
The algorithm parameters were set at $\gamma=0$ (leading to opportunistic behaviour), $\alpha=0.05$, having linearly annealed $\epsilon=0.2 \rightarrow 0$ by $t=3000$.
Benign and malicious hosts uploaded between \SIrange{0}{1}{\mega\bit\per\second} and \SIrange{2.5}{6}{\mega\bit\per\second} respectively, and hosts were redrawn at each episode's start with $\operatorname{P}(\mathit{malicious})=0.4$.
$U_s$ was fixed at $klmn+2$ \si{\mega\bit\per\second}.
The performance of each choice of $n$ was averaged over \num{10} episodes of length \num{10000} timesteps (setting each agent's $\wvec{}=\bm{0}$ between episodes).
Host allocations were generated pseudorandomly to ensure fairness between choices of $n$.

@@

\Cref{fig:marl-granularity} shows that allocating more hosts to each learner leads to a guaranteed decrease in the amount of benign traffic observed at the server.
?? TODO: rephrase.
This manifests in two ways: the best-achievable performance decreases, as does the learning rate.
Given the UDP-like behaviour of Tcpreplay, being able to replicate the online performance of \citeauthor{DBLP:journals/eaai/MalialisK15} in the $n=2$ case suggests that the numeric simulation their work is entirely based upon does not capture the key interactions between TCP flows and packet drop as a control action.
?? Expand according to the summary you posted in the notes document.

\subsection{TCP Multiplexing and Fallback Behaviour}
While Tcpreplay was useful in recreating the original work and examining the above condition, flow behaviour under packet loss is closer to that of UDP when using it to replay traffic.
To investigate the effects of collateral damage upon benign TCP flows, an iperf3 test was conducted using mininet.
Several hosts communicated concurrently to a single server over one switch, with maximum send rates in $\{1,...,5\}$ \si{\mega\bit\per\second} and a constant \SI{40}{\milli\second} latency.
During the test, the rate of packet drop was increased linearly from \SIrange{0}{40}{\percent} and the observed throughput for each host was measured.

\begin{figure}
	\includegraphics[width=\linewidth]{../plots/mplex}
	\caption{
		Average TCP upload rate of hosts targeting different bandwidths in \emph{mininet}, computed over 10 runs.
		As the rate of packet drop is linearly increased from \SIrange{0}{40}{\percent}, all flows begin to converge on an extremely low upload rate, regardless of the target bandwidth.
		This behaviour is expected and well-understood, and is present in this emulated environment---prior treatments have ignored such phenomena, which have strong implications on agent and network design.
		Malicious flows are not expected to obey the Mathis equation in this way, worsening collateral damage experienced by benign TCP flows.
		\label{fig:mplex-tcp}
	}
\end{figure}

\Cref{fig:mplex-tcp} shows that benign TCP flows are severely harmed by \emph{any} packet drop, so collateral damage to these hosts is more significant.
Good-faith TCP congestion avoidance causes these flows to attempt to back off very quickly.
%?? Talk about the same major points you made in the presentation here (i.e. TCP interactions) w/ pushback. Or in the evaluation...
UDP, however, behaves as expected---malicious hosts are expected to exhibit the same behaviour, given that congestion is their primary goal.

@@

@@

%?? Abuses of RL 
Earnest, well-considered application of RL towards the challenge of intrusion detection/prevention has seen comparatively little examination.
Past work exhibits abuses of the paradigm as a glorified classifier for anomaly detection \cite{shamshirband2014anomaly} and DDoS prevention \cite{DBLP:conf/mates/ServinK08}.
?? Explain/be less incendiary...

For categorising how RL fits into solving problems, we label works as direct- or indirect-control RL.
A \emph{direct-control} RL problem is one where the RL agent(s) are learning optimal control over a set of actions to act as the \emph{primary} defence or decision-maker---requiring measurements, reward functions and action sets tailored for this purpose.
?? Good Direct-control work? What this work is based on (which directly captures the spirit of ) \cite{DBLP:phd/ethos/Malialis14, DBLP:journals/eaai/MalialisK15}

An \emph{indirect-control} RL problem, where the role played by the agents is to act in service to \emph{another technique} responsible for decision-making, further optimising or generalising aspects of its operation beyond that of hand-coded heuristics.
%Indirect-control applications: link to all the juicy stuff in e.g. HotNets, data-driven networking, the works. \cite{DBLP:conf/hotnets/ValadarskySST17, DBLP:conf/hotnets/MaoAMK16}.
A past example includes learning when best to \emph{communicate} and share knowledge between \emph{hidden Markov model} anomaly detectors \cite{DBLP:conf/paisi/XuSH07}.
The position of this work is weakened by its reliance on the discredited `DARPA99' dataset \cite{DARPA-IDD, DBLP:conf/cisda/TavallaeeBLG09, DBLP:conf/sp/SommerP10}, but the idea itself is well-treated and this acts as a driver for improvements in this direction.
Outside of anomaly/intrusion detection, there has been growing interest in the use of reinforcement learning in data-driven networking, such as for intra-AS route optimisation \cite{DBLP:conf/hotnets/ValadarskySST17} and for resource-constrained process allocation \cite{DBLP:conf/hotnets/MaoAMK16}.
These works emphasise the necessity of ingenuity in effectively handling how states and actions are represented.

>>>>

?? Probably want to talk about NIDS/IPS, DDoS attack variants (leading into characteristics, supporting features). Amplification (UDP \cite{DBLP:conf/ndss/Rossow14}, TCP \cite{DBLP:conf/uss/KuhrerHRH14}), Transit-link (Crossfire \cite{DBLP:conf/sp/KangLG13}, Coremelt \cite{DBLP:conf/esorics/StuderP09}). Mirai botnet's involvement \cite{DBLP:conf/uss/AntonakakisABBB17}.

?? What makes RL a suitable method for network anomaly detection, what features are most relevant?
?? Point I was thinking of: feedback-loop-like model allows monitoring \emph{after} an action is taken to (in theory) allow forgiveness of mistakenly punished flows. This does hinge on taking a flow-by-flow look at the state space, but if we can combine knowledge of current state (duh!), the last action taken (i.e. an indicator of our previous assessment [such as high pdrop $\implies$ bad]) then perhaps a flow which falls off identically to a legit flow can be rescued.

?? Relevant features: aggregate network state (load at various points [this has been done, of course]), flow-specific measurements (upload/download ratio when bandwidth above threshold \cite{DBLP:conf/ndss/Rossow14}, packet inter-arrival times, etc.)
?? Need to find a good survey which suggests indicative features.
?? If we assume amplification attacks, we know it won't be `random' source IPs (since it's mostly-legit servers who think that they're doing a good job by replying)

@@

To test this, we configured the network topology using $k=2$ teams, $l=3$ intermediate nodes per team, $m=2$ agents per intermediate node, and $n \in \{2, 4, 8, 16\}$ hosts per learner.
This is a slight simplification of the \textquote{online} experiment as presented by \textcite{DBLP:journals/eaai/MalialisK15}, choosing fewer teams but remaining as a single-server with a fan-out network.
The algorithm parameters were set at $\gamma=0$ (leading to opportunistic behaviour), $\alpha=0.05$, having linearly annealed $\epsilon=0.2 \rightarrow 0$ by $t=3000$.
Benign and malicious hosts uploaded between \SIrange{0}{1}{\mega\bit\per\second} and \SIrange{2.5}{6}{\mega\bit\per\second} respectively, and hosts were redrawn at each episode's start with $\operatorname{P}(\mathit{malicious})=0.4$.
$U_s$ was fixed at $klmn+2$ \si{\mega\bit\per\second}.
The performance of each choice of $n$ was averaged over \num{10} episodes of length \num{10000} timesteps (setting each agent's $\wvec{}=\bm{0}$ between episodes).
Host allocations were generated pseudorandomly to ensure fairness between choices of $n$.

@@

\Cref{fig:marl-granularity} shows that as more hosts are allocated to each learner, the fraction of traffic believed to be good (as observed at the server) decreases.
This manifests in two ways: the best-achievable performance decreases, as does the learning rate.
Given the UDP-like behaviour of Tcpreplay, being able to replicate the online performance of \citeauthor{DBLP:journals/eaai/MalialisK15} in the $n=2$ case suggests that the numeric simulation their work is entirely based upon does not capture the key interactions between TCP flows and packet drop as a control action.
From the findings of \cref{sec:tcp-multiplexing-and-fallback-behaviour} and the Mathis equation, their results should be considerably worse in practice.
%?? Expand according to the summary you posted in the notes document.
%That the use of UDP-like traffic replicates their experiments, compared against how real tcp behaves in mininet and according to the Mathis equation, shows that they really did not consider the behaviour of benign TCP in this environ (i.e., their results should be considerably worse in practice).

\subsection{TCP Multiplexing and Fallback Behaviour}\label{sec:tcp-multiplexing-and-fallback-behaviour}
While Tcpreplay was useful in recreating the original work and examining the above condition, flow behaviour under packet loss is closer to that of UDP when using it to replay traffic.
To investigate the effects of collateral damage upon benign TCP flows, an iperf3 test was conducted using mininet.
Several hosts communicated concurrently to a single server over one switch, with maximum send rates in $\{1,...,5\}$ \si{\mega\bit\per\second} and a constant \SI{40}{\milli\second} latency.
The rate of packet drop was increased linearly from \SIrange{0}{40}{\percent} and the observed throughput for each host was measured, to show a trend and determine for the mininet environment whether throughput would decrease as expected, at a super-linear rate.

\begin{figure}
	\includegraphics[width=\linewidth]{../plots/mplex}
	\caption{
		Average TCP upload rate of hosts targeting different bandwidths in \emph{mininet}, computed over 10 runs, for choices of packet drop rate up until throughput is reduced to almost \SI{0}{\mega\bit\per\second}.
		As the rate of packet drop increases from \SIrange{0}{40}{\percent}, all flows converge on an extremely low throughput, regardless of the target bandwidth---a super-linear decrease for legitimate TCP traffic.
		Prior treatments have ignored such phenomena, which have strong implications on agent and network design.
		Malicious flows are not expected to obey the Mathis equation in this way, worsening collateral damage experienced by benign TCP flows.
		\label{fig:mplex-tcp}
	}
\end{figure}

\Cref{fig:mplex-tcp} shows that this trend is present.
Accordingly, benign TCP flows are severely harmed by \emph{any} packet drop, so collateral damage to these hosts is more significant.
Good-faith TCP congestion avoidance causes these flows to attempt to back off very quickly.
%?? Talk about the same major points you made in the presentation here (i.e. TCP interactions) w/ pushback. Or in the evaluation...
UDP, however, behaves as expected---malicious hosts are expected to exhibit the same behaviour, given that congestion is their primary goal.

@@

?? Existing sims used for testing such applications reliant on traces, or not sophisticated enough to capture interactive, back-and-forth (correlated) behaviours---possibly discarded as second-hand effects by past work when these are so crucial given user traffic patterns (and the nature of the control signal we choose to enact).

@@

%?? Abuses of RL 
Earnest, well-considered application of RL towards the challenge of intrusion detection/prevention has seen comparatively little examination.
Past work exhibits treatment of the paradigm as a traditional classifier for anomaly detection \cite{shamshirband2014anomaly} and DDoS prevention \cite{DBLP:conf/mates/ServinK08}.
Given that arguably the main strength of RL techniques is their ability to control ongoing interaction and to adapt by observing the concrete effects of actions taken, such works fail to see how best to apply the rich literature on the subject.

For categorising how RL fits into solving problems, we label works as direct- or indirect-control RL.
A \emph{direct-control} RL problem is one where the RL agent(s) are learning optimal control over a set of actions to act as the \emph{primary} defence or decision-maker---requiring measurements, reward functions and action sets tailored for this purpose.
?? Good Direct-control work? What this work is based on (which directly captures the spirit of RL) \cite{DBLP:phd/ethos/Malialis14, DBLP:journals/eaai/MalialisK15}

An \emph{indirect-control} RL problem, where the role played by the agents is to act in service to \emph{another technique} responsible for decision-making, further optimising or generalising aspects of its operation beyond that of hand-coded heuristics.
%Indirect-control applications: link to all the juicy stuff in e.g. HotNets, data-driven networking, the works. \cite{DBLP:conf/hotnets/ValadarskySST17, DBLP:conf/hotnets/MaoAMK16}.
A past example includes learning when best to \emph{communicate} and share knowledge between \emph{hidden Markov model} anomaly detectors \cite{DBLP:conf/paisi/XuSH07}.
The position of this work is weakened by its reliance on the discredited `DARPA99' dataset \cite{DARPA-IDD, DBLP:conf/cisda/TavallaeeBLG09, DBLP:conf/sp/SommerP10}, but the idea itself is well-treated and this acts as a driver for improvements in this direction.
Outside of anomaly/intrusion detection, there has been growing interest in the use of reinforcement learning in data-driven networking, such as for intra-AS route optimisation \cite{DBLP:conf/hotnets/ValadarskySST17} and for resource-constrained process allocation \cite{DBLP:conf/hotnets/MaoAMK16}.
?? Also adaptively choosing video bitrates based on network conditions \cite{DBLP:conf/sigcomm/MaoNA17}.
These works emphasise the necessity of ingenuity in effectively handling how states and actions are represented.
